Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-05-22 awipev-CO2.Rmd
- Heavily edited and corrected chunk "read and prepare discrete data" - found an error in nyalesund_cc

```{r Read, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_corr = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  )
d_all <- d_hour #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE))
#restore datetime, date which are mangled by the murtates above
d_all$datetime <- d_hour$datetime
d_all$date <- d_hour$date
d_all$month <- month(d_all$datetime)
```

@@

```{r Temp filter, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}
# Binding temp_insitu_11m + temp_insitu_ctd if we get gaps in temp_insitu_11m = call this new column = temp_mix 
d_all <- d_all %>%
  dplyr::mutate(temp_mix = ifelse(!is.na(temp_insitu_11m), temp_insitu_11m, temp_insitu_ctd_9m))
```

```{r read and prepare discrete data, echo=FALSE, warning=FALSE, message=FALSE}
d <- tibble(
  read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
)
#d$datetime <- as.POSIXct(d$sampling_date, format="%d/%m/%Y %H:%M", tz="UTC")
d$datetime <- dmy_hm(d$sampling_date, tz="UTC")
d$measure_date <- dmy(d$measure_date, tz="UTC")
d$date <- as.Date(d$datetime)

@@

# Keep only flag = 2 for at/ct analysis 
# Keep only flag = 2 for pH durafet and seaFET.
d <- d%>%
  dplyr::mutate(ct= replace(ct, qflag_ct != 2, NA),
                at= replace(at, qflag_at != 2, NA),
                pH_s_seafet= replace(pH_s_seafet, qflag_pH_s_seafet != 2, NA),
                pH_s_durafet= replace(pH_s_durafet, qflag_pH_s_durafet != 2, NA))

@@

# mean at and ct
at_ct_mean <- d %>% 
#  group_by(datetime)%>%
  group_by(date) %>%
  dplyr::summarize(at_mean = mean(at, na.rm = T),
                   ct_mean = mean(ct, na.rm = T),
                   at_sd = sd(at, na.rm = T), 
                   ct_sd = sd(ct, na.rm = T)
                   )
d <- left_join(d, at_ct_mean, by='date') %>% 
  dplyr::select(datetime, date, temp_s_lab, pH_s_seafet, pH_s_durafet, location_flag, at_mean, ct_mean) %>% # removed useless variables and also to do unique next
  dplyr::distinct()

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>% dplyr::select(c(closest_datetime, sal_fb, temp_fb, temp_mix, sal_mix, temp_insitu_ctd, pressure_insitu_ctd)), by="closest_datetime")

# Conversion of spectro pH to insitu mix T°
pH_seafet <- d %>%
  dplyr::filter(!is.na(pH_s_seafet) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>%
  dplyr::mutate(ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$pH_s_seafet, ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd, S = pH_seafet$sal_mix, Pt = 0, Sit = 0))

pH_durafet <- d %>%
#  dplyr::filter(!is.na(pH_s_durafet))
    dplyr::filter(!is.na(pH_s_durafet) & !is.na(sal_fb) & !is.na(temp_fb))
pH_durafet <- pH_durafet %>%
#  dplyr::mutate(ph_s_dur_temp_fb = pHinsi(pH=pH_durafet$pH_s_durafet, ALK=pH_durafet$at_mean_interp*1e-6, Tinsi=pH_durafet$temp_fb, Tlab=pH_durafet$temp_s_lab, Pinsi= pH_durafet$depth/10, S=pH_durafet$sal_fb, Pt=0, Sit=0))
  dplyr::mutate(ph_s_dur_temp_fb = pHinsi(pH=pH_durafet$pH_s_durafet, ALK=pH_durafet$at_mean_interp*1e-6, Tinsi=pH_durafet$temp_fb, Tlab=pH_durafet$temp_s_lab, Pinsi= 0, S=pH_durafet$sal_fb, Pt=0, Sit=0))

@@

#  make means in "d" to have same format and call it "discrete"
discrete <- d %>% 
  dplyr::group_by(datetime) %>% 
  dplyr::summarize(sal_fb = mean(sal_fb, na.rm = T),
                   temp_fb = mean(temp_fb, na.rm = T),
                   at_mean_interp = mean(at_mean_interp, na.rm = T),
            at = mean(at, na.rm = T),
            qflag_at = mean(qflag_at, na.rm = T),
            ct_mean_interp = mean(ct_mean_interp, na.rm = T),
            ct = mean(ct, na.rm = T),
            qflag_ct = mean(qflag_ct, na.rm = T)
            )
# join discrete + seafet/durafet
discrete <- left_join(discrete, pH_seafet_mean, by='datetime')
discrete <- left_join(discrete, pH_durafet_mean, by='datetime')

##   *****Attention, temp noted by logisticians in d is FB temp with SBE45 ******
##   *****Attention, I asked to the logisticians to log insitu temp from 9 june 2019

discrete_no_na <- discrete %>%
  dplyr::filter(!is.na(sal_fb)  & !is.na(temp_fb) & !is.na(at) & !is.na(ct))
discrete_is_na <- discrete %>%
  dplyr::filter(is.na(sal_fb)  | is.na(temp_fb) | is.na(at) | is.na(ct))

carb <- carb(15, discrete_no_na$at*1e-6, discrete_no_na$ct*1e-6, S = discrete_no_na$sal_fb, T = discrete_no_na$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")

discrete_no_na$phcalc<- carb$pH
discrete_no_na$pco2_calc<- carb$pCO2

#create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$phcalc<- NA
discrete_is_na$pco2_calc<- NA
discrete <- rbind(discrete_no_na, discrete_is_na)
discrete <-discrete%>%dplyr::arrange(datetime)
```

@@

# make the mean of E0 by period of deployement
calib_spectro <- group_by(calib_spectro, deployment_sf) %>%
  summarise(E0int25 = mean(E0int25, na.rm = T),
            E0ext25 = mean(E0ext25, na.rm = T) 
            )
names(calib_spectro) <- c("deployment_sf", "E0int25_sf", "E0ext25_sf")

>>>>

```{r Read, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_corr = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  )
d_all <- d_hour #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE))
#restore datetime, date which are mangled by the mutates above
d_all$datetime <- d_hour$datetime
d_all$date <- d_hour$date
d_all$month <- month(d_all$datetime)
```

@@

```{r Temp filter, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}
# Create temp_mix: filling gaps in temp_insitu_11m using temp_insitu_9m
d_all <- d_all %>%
  dplyr::mutate(temp_mix = ifelse(!is.na(temp_insitu_11m), temp_insitu_11m, temp_insitu_ctd_9m))
```

```{r read and prepare discrete data, echo=FALSE, warning=FALSE, message=FALSE}
d <- tibble(
  read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
) %>% 
  dplyr::mutate(datetime = dmy_hm(sampling_date, tz="UTC"),
                measure_date = dmy(measure_date, tz="UTC"),
                date = as.Date(datetime)) %>% 
  dplyr::rename(ph_s_sf = pH_s_seafet,
                ph_s_dur = pH_s_durafet,
                qflag_ph_s_sf = qflag_pH_s_seafet,
                qflag_ph_s_dur = qflag_pH_s_durafet)

@@

# Keep only flag = 2 for at/ct analysis 
# Keep only flag = 2 for pH durafet and seaFET.
d <- d%>%
  dplyr::mutate(ct = replace(ct, qflag_ct != 2, NA),
                at = replace(at, qflag_at != 2, NA),
                ph_s_sf = replace(ph_s_sf, qflag_ph_s_sf != 2, NA),
                ph_s_dur = replace(ph_s_dur, qflag_ph_s_dur != 2, NA))

@@

# mean at and ct
at_ct_mean <- d %>% 
#  group_by(datetime)%>%
  group_by(date) %>%
  dplyr::summarize(at_mean = mean(at, na.rm = TRUE),
                   ct_mean = mean(ct, na.rm = TRUE),
                   at_sd = sd(at, na.rm = TRUE), 
                   ct_sd = sd(ct, na.rm = TRUE)
                   )
d <- left_join(d, at_ct_mean, by='date') %>% 
  dplyr::select(datetime, date, temp_s_lab, ph_s_sf, ph_s_dur, location_flag, at, ct, at_mean, ct_mean) %>% # removed useless variables and also to do unique next
  dplyr::distinct()

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(
                     closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     pressure_insitu_ctd
                   )
                 ), by = "closest_datetime")

# Conversion of spec pH of in situ samples to in situ temp
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, S = pH_seafet$sal_mix, Pt = 0, Sit = 0))

# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  dplyr::filter(!is.na(ph_s_dur) & !is.na(sal_fb) & !is.na(temp_fb))
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, Pt = 0, Sit = 0))

@@

#  make means in "d" to have same format and call it "discrete"
discrete <- d %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    sal_fb = mean(sal_fb, na.rm = TRUE),
    sal_mix = mean(sal_mix, na.rm = TRUE),
    temp_fb = mean(temp_fb, na.rm = TRUE),
    at_mean_interp = mean(at_mean_interp, na.rm = TRUE),
    at = mean(at, na.rm = TRUE),
    #qflag_at = mean(qflag_at, na.rm = TRUE), # already dealt with above
    ct_mean_interp = mean(ct_mean_interp, na.rm = TRUE),
    ct = mean(ct, na.rm = TRUE),
    #qflag_ct = mean(qflag_ct, na.rm = TRUE)  # already dealt with above
    temp_mix = mean(temp_mix, na.rm = TRUE),
    temp_insitu_11m = mean(temp_insitu_11m, na.rm = TRUE)
  )
# join discrete + seafet/durafet
discrete <- left_join(discrete, pH_seafet_mean, by='date')
discrete <- left_join(discrete, pH_durafet_mean, by='date')

##   *****Attention, temp noted by logisticians in d is FB temp with SBE45 ****** $SA$
##   *****Attention, I asked to the logisticians to log insitu temp from 9 june 2019

# pH calculated (at and ct) at Ferrybox sal and temp
discrete_no_na <- discrete %>%
  dplyr::filter(!is.na(sal_fb)  & !is.na(temp_fb) & !is.na(at) & !is.na(ct))
discrete_is_na <- discrete %>%
  dplyr::filter(is.na(sal_fb)  | is.na(temp_fb) | is.na(at) | is.na(ct))
carb <- carb(15, discrete_no_na$at*1e-6, discrete_no_na$ct*1e-6, S = discrete_no_na$sal_fb, T = discrete_no_na$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
discrete_no_na$phcalc_fb <- carb$pH # pH to compare with durafet pH (at Ferrybox temp and sal)
#discrete_no_na$pco2_calc<- carb$pCO2

# create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$phcalc_fb <- NA
#discrete_is_na$pco2_calc<- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)

# pH calculated (at and ct) at in situ (11 m) sal and temp
discrete_no_na <- discrete %>%
  dplyr::filter(!is.na(sal_mix)  & !is.na(temp_insitu_11m) & !is.na(at) & !is.na(ct))
discrete_is_na <- discrete %>%
  dplyr::filter(is.na(sal_mix)  | is.na(temp_insitu_11m) | is.na(at) | is.na(ct))
carb <- carb(15, discrete_no_na$at*1e-6, discrete_no_na$ct*1e-6, S = discrete_no_na$sal_mix, T = discrete_no_na$temp_insitu_11m, P = 11/10, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
discrete_no_na$phcalc_insitu <- carb$pH # pH to compare with durafet pH (at Ferrybox temp and sal)
discrete_no_na$pco2_calc_insitu <- carb$pCO2

#create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$phcalc_insitu <- NA
discrete_no_na$pco2_calc_insitu <- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)

@@

# make the mean of E0 by period of deployement
calib_spectro <- group_by(calib_spectro, deployment_sf) %>%
  summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            )
names(calib_spectro) <- c("deployment_sf", "E0int25_sf", "E0ext25_sf")
